---
title: "Get started with spinviz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with spinviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

```{r}
library(spinviz)
```

`spinviz` produces two kinds of injury visualisations:

- **Heatmaps** of injury frequency projected onto human-body diagrams (`heatmap_diagram()`), with front, back, or both views and male or female body templates.
- **Interactive sunburst diagrams** of tissue/pathology classifications (`sunburst_diagram_echarts()`), built on echarts4r.

## Data format

The package expects a data frame with two descriptor columns followed by one or more columns of injury frequencies (one per sport or cohort).

### Heatmap data

| region_area | subcategory | sport1 |
|---|---|---|
| Head and Neck | Head | 10 |
| Head and Neck | Neck | 20 |
| Upper Limb | Shoulder | 5 |
| Upper Limb | Upper Arm | 8 |

The 18 recognised body subcategories and their `region_area` groupings, plus a special `Unspecified` category for injuries that cannot be localised, are available as the built-in `body_categories` taxonomy (19 rows).

### Sunburst data

For the sunburst, the first two columns are `tissue` and `pathology`:

| tissue | pathology | sport1 |
|---|---|---|
| Muscle / Tendon | Muscle strain | 20 |
| Muscle / Tendon | Tendon rupture | 10 |
| Bone | Fracture | 16 |
| Bone | Bone contusion | 6 |

The built-in `injury_categories` taxonomy contains the 25-row tissue/pathology classification.

## Heatmap example

```{r}
subcategory <- c("Head", "Neck", "Shoulder", "Chest", "Upper Arm", "Elbow",
                 "Abdomen", "Forearm", "Hip Groin", "Wrist", "Hand",
                 "Thigh", "Knee", "Lower Leg", "Ankle", "Foot",
                 "Thoracic Spine", "Lumbosacral")

region_area <- rep("Example", length(subcategory))
boxing <- c(15, 5, 18, 12, 20, 6, 10, 14, 9, 9, 11, 3, 16, 13, 7, 8, 18, 22)

df <- data.frame(region_area, subcategory, boxing)
```

Then call `heatmap_diagram(injury_data, selected_sport, view_choice)`, where
`view_choice` is one of `"front"`, `"back"`, or `"both"`:

```{r}
#| fig.width: 6
#| fig.height: 8
heatmap_diagram(df, "boxing", "front", sex = "male", show_scale = FALSE)
```

See `?heatmap_diagram` for options controlling labels, values, opacity, the
legend scale, and custom palettes. Rows with an `"Unspecified"` subcategory
are shown as a label below the diagram (they have no body region to colour).

## Sunburst example

The quickest way to get a correctly-ordered taxonomy is `injury_categories`:

```{r}
df_sb <- injury_categories
df_sb$boxing <- c(20, 0, 0, 10, 31, 18, 16, 16, 20, 20, 27, 46,
                  67, 31, 54, 20, 27, 30, 96, 82, 48, 26, 33, 34, 24)

sunburst_diagram_echarts(df_sb, "boxing", plot_title = "Boxing Injuries")
```

The sunburst is interactive: tissue types sit on the inner ring, pathologies
on the outer ring. See `?sunburst_diagram_echarts` for options such as
`depth = 1` (tissue ring only), excluding "Unspecified"/"Non-specific" rows,
and label/radius/font tuning.

## Convenience wrappers

If you use the package's standard classifications, you can skip constructing
the descriptor columns entirely and pass just a vector of counts:

```{r}
#| fig.width: 6
#| fig.height: 8
heatmap_diagram_default(c(boxing, unspecified = 4), "front")
```

```{r}
sunburst_diagram_default(df_sb$boxing, plot_title = "Boxing Injuries")
```

Counts are matched by row position to `body_categories` / `injury_categories`.

## Next steps

- `vignette("importing-csv-data", package = "spinviz")` — a template-based CSV workflow
- `vignette("customisation", package = "spinviz")` — palettes, saving diagrams, and display options
