---
title: "Customising and saving diagrams"
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Colour palettes

Both diagram functions accept a palette as either:

- A palette name supported by `diagram_colours()` (viridis or HCL palettes),
  e.g. `"magma"`, `"plasma"`, `"Reds"`
- A character vector of hex colours,
  e.g. `c("#FFFFFF", "#FFFF00", "#FF0000")`

Run `?diagram_colours` to see supported palette names, and use `test_colour()`
to visualise a named or custom palette before applying it.

```{r}
test_colour("magma")
test_colour(c("#FFFFFF", "#FFFF00", "#FF0000"))
```

## Heatmap display options

`heatmap_diagram()` supports:

- View: `"front"`, `"back"`, or `"both"`
- Diagram sex: `"male"` or `"female"`
- Opacity control
- Optional labels (region names) and/or values
- Optional legend scale

```{r}
heatmap_diagram(df, "boxing", "both", sex = "female",
                opacity = 0.8, show_scale = TRUE)
```

See `?heatmap_diagram` for the full argument list.

## Sunburst display options

`sunburst_diagram_echarts()` supports:

- Colour per tissue type, with pathology slices automatically lightened
- Optional exclusion of "Unspecified"/"Non-specific" tissue rows
- `depth = 1` to show only the tissue-level ring
- Label/leader-line, radius, font and title tuning

```{r}
sunburst_diagram_echarts(df_sb, "boxing", depth = 1,
                         plot_title = "Tissue types only")
```

See `?sunburst_diagram_echarts` for the full argument list.

## Saving diagrams

`save_diagram(plot, file, width = ...)` detects whether `plot` is a heatmap or
a sunburst and applies the matching export logic, deriving the height from the
width at the correct aspect ratio automatically:

```{r}
p_heat <- heatmap_diagram(df, "boxing", "front", sex = "male")
save_diagram(p_heat, file = "heatmap.png", width = 1600, units = "px")

p_sun <- sunburst_diagram_echarts(df_sb, "boxing")
save_diagram(p_sun, file = "sunburst.png", width = 1200)
```

Heatmaps are saved via `ggplot2::ggsave()` (PNG, PDF, SVG, JPG, ...). Sunbursts
are saved via a headless Chromium browser through the optional `chromote`
package (PNG, JPG, PDF), so install `chromote` and `base64enc` if you need
sunburst file export.
