---
title: "Importing data from a CSV file"
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

If your injury data lives in a CSV file, `spinviz` provides helpers for the
full workflow: create a template, fill it in, read it back, and plot.

## Heatmap (body injury) data

**1. Create a template CSV** with all 18 recognised body subcategories and one
empty column per sport:

```{r}
create_injury_template("injuries.csv", sports = c("boxing", "judo"))
```

**2. Fill in the file** in your favourite spreadsheet editor, entering the
injury frequencies and the `region_area` grouping for each row.

**3. Read the file back in.** `read_injury_data()` checks that the required
columns are present, that at least one sport column exists, and that sport
values are numeric (with a warning for values that fail to parse):

```{r}
df <- read_injury_data("injuries.csv")
```

**4. Plot as usual:**

```{r}
heatmap_diagram(df, "boxing", "front", sex = "male")
```

## Sunburst (tissue/pathology) data

The same workflow exists for the sunburst format. `create_sunburst_template()`
writes a CSV pre-filled with the 25-row `injury_categories` taxonomy (`tissue`
and `pathology` columns), and `read_sunburst_data()` validates it — including
filling blank `tissue` cells down from the row above, so compact hand-edited
files work:

```{r}
create_sunburst_template("tissue_injuries.csv", sports = c("boxing", "judo"))
df <- read_sunburst_data("tissue_injuries.csv")
sunburst_diagram_echarts(df, "boxing")
```
